---
title: "p8105_hw3_qs2200"
author: "Qi Shao"
date: "10/7/2018"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(ggridges)
library(patchwork)
library(p8105.datasets)
```

# Problem 1

**1. Import and clean the dataset.**
```{r problem 1.1}
brfss_df = p8105.datasets::brfss_smart2010 %>%
  janitor::clean_names() %>%
  filter(topic == "Overall Health") %>%
  mutate(response = factor(response, levels = c("Excellent", "Very good", "Good","Fair", "Poor")))%>%
  rename(state = locationabbr) %>%
  arrange(response) 

##filter(topic == "Overall Health", response ==c("Excellent", "Very good", "Good","Fair", "Poor"))
##why can't we filter the rows which response is excellent to poor
```
**2. In 2002, which states were observed at 7 locations?**
```{r problem 1.2}
brfss_state = 
  group_by(brfss_df , state) %>%
  summarize(n = n_distinct(locationdesc)) %>%
  filter(n == 7)
```

The 7 states `r brfss_state$state` were observed at 7 locations.

**3. Make a “spaghetti plot” that shows the number of observations in each state from 2002 to 2010.**
```{r problem 1.3}

brfss_spa =
  group_by(brfss_df, state,year) %>%
  summarize(n_obs = n())

ggplot(brfss_spa, aes(x = year, y = n_obs, color = state))+
  geom_line()+geom_point()+
  labs(x = "Year",
    y = "Number of observations")

```
**4. Make a table showing, for the years 2002, 2006, and 2010, the mean and standard deviation of the proportion of “Excellent” responses across locations in NY State.**

```{r problem 1.4}

brfss_tb =
  filter(brfss_df, year %in% c(2002, 2006, 2010),
         state == "NY",
         response == "Excellent")%>%
  group_by(year) %>%
  summarize(mean_prop = mean(data_value), 
            sd_prop = sd(data_value, na.rm  = TRUE))
  knitr::kable(brfss_tb, digits = 2)

```


**5.For each year and state, compute the average proportion in each response category (taking the average across locations in a state). Make a five-panel plot that shows, for each response category separately, the distribution of these state-level averages over time.**

```{r problem 1.5, message = FALSE}
aver_exce = 
  filter(brfss_df, response == "Excellent") %>%
  group_by(state,year) %>%
  summarize(mean_prop = mean(data_value)) %>%
  ggplot(aes(x = year, y = mean_prop, color = state)) +
  geom_line() +
  theme(legend.position = "none")

aver_very = 
  filter(brfss_df, response == "Very good") %>%
  group_by(state,year) %>%
  summarize(mean_prop = mean(data_value)) %>%
  ggplot(aes(x = year, y = mean_prop, color = state)) +
  geom_line() +
  theme(legend.position = "none")

aver_good = 
  filter(brfss_df, response == "Good") %>%
  group_by(state,year) %>%
  summarize(mean_prop = mean(data_value)) %>%
  ggplot(aes(x = year, y = mean_prop, color = state)) +
  geom_line() +
  theme(legend.position = "none")

aver_fair = 
  filter(brfss_df, response == "Fair") %>%
  group_by(state,year) %>%
  summarize(mean_prop = mean(data_value)) %>%
  ggplot(aes(x = year, y = mean_prop, color = state)) +
  geom_line() +
  theme(legend.position = "none")

aver_poor = 
  filter(brfss_df, response == "Poor") %>%
  group_by(state,year) %>%
  summarize(mean_prop = mean(data_value)) %>%
  ggplot(aes(x = year, y = mean_prop, color = state)) +
  geom_line() +
  theme(legend.position = "bottom")

aver_exce + aver_very + aver_good + aver_fair + aver_poor
  
```


## Problem 2 

**1. Import, explore and describe the dataset.**
```{r problem 2.1}
instacart_df = p8105.datasets::instacart 
head(instacart_df)
```

The dataset is about online purchase order in Instacart. 

There are `r ncol(instacart_df)` variables in the dataset, in which the "eval_set", "aisle", "product_name", "department" are character variables, others are all integer variables.These variables contain the id and category of products in each order, the user and id of an order, the time each order is placed and the time each product is added to the cart and so on.

There are `r nrow(instacart_df)` observations in the dataset, each one represents a product in an order. For example, the first row in the dataset illustrate the order id is 1, the product id is 49302. It is the first product added in cart in order 1, and it has been reordered by the user 112108 in the past.This order belongs to the train evaluation set, and is the 4th order of the user. This order is placed on the 4th day of a week, 10 a.m., and it has been 9 days since last order was placed.The product is Bulgarian Yogurt, in dairy and egg department (id: 16), yogurt aisle(id: 120).

**2. How many aisles are there, and which aisles are the most items ordered from?**
```{r problem 2.2}
aisle_max = as.data.frame(table(instacart_df$aisle))%>%
  filter(Freq == max(Freq)) 
```

There are `r length(unique(instacart_df$aisle))` alsles there.The aisle "`r aisle_max$Var1`" is the most items ordered from.

**3. Make a plot that shows the number of items ordered in each aisle. Order aisles sensibly, and organize your plot so others can read it.**
```{r problem 2.3}
instacart_df %>%
  group_by(aisle, department) %>%
  summarize(n = n()) %>%
  ggplot(aes(x = aisle, y = n))+
  geom_bar(stat="identity")+
  facet_wrap( ~ department,ncol = 3,scale = "free")

```

**4. Make a table showing the most popular item aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”.**
```{r problem 2.4}
instacart_df %>%
  group_by(aisle, product_name) %>%
  summarize(item_number = n()) %>%
  filter(aisle %in% c("baking ingredients", "dog food care",  "packaged vegetables fruits")) %>%
  filter(item_number == max(item_number)) %>%
  knitr::kable(digits = 0)
```

**5. Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers.** (i.e. produce a 2 x 7 table)
```{r problem 2.5}
instacart_df %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(order_dow, product_name) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  spread(key = product_name, value = mean_hour) %>%
  knitr::kable(digits = 1)
```
  

```