---
title: "p8105_hw3_qs2200"
author: "Qi Shao"
date: "10/7/2018"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(ggridges)
library(patchwork)
library(p8105.datasets)
```

# Problem 1

**1. Import and clean the dataset.**
```{r problem 1.1}
brfss_df = p8105.datasets::brfss_smart2010 %>%
  janitor::clean_names() %>%
  filter(topic == "Overall Health") %>%
  mutate(response = factor(response, levels = c("Excellent", "Very good", "Good","Fair", "Poor")))%>%
  rename(state = locationabbr) %>%
  arrange(response) 

##filter(topic == "Overall Health", response ==c("Excellent", "Very good", "Good","Fair", "Poor"))
##why can't we filter the rows which response is excellent to poor
```
**2. In 2002, which states were observed at 7 locations?**
```{r problem 1.2}
brfss_state = 
  group_by(brfss_df , state) %>%
  summarize(n = n_distinct(locationdesc)) %>%
  filter(n == 7)
```

The 7 states `r brfss_state$state` were observed at 7 locations.

**3. Make a “spaghetti plot” that shows the number of observations in each state from 2002 to 2010.**
```{r problem 1.3}

brfss_spa =
  group_by(brfss_df, state,year) %>%
  summarize(n_obs = n())

ggplot(brfss_spa, aes(x = year, y = n_obs, color = state))+
  geom_line()+geom_point()+
  labs(x = "Year",
    y = "Number of observations")

```
**4. Make a table showing, for the years 2002, 2006, and 2010, the mean and standard deviation of the proportion of “Excellent” responses across locations in NY State.**

```{r problem 1.4}

brfss_tb =
  filter(brfss_df, year %in% c(2002, 2006, 2010),
         state == "NY",
         response == "Excellent")%>%
  group_by(year) %>%
  summarize(mean_prop = mean(data_value), 
            sd_prop = sd(data_value, na.rm  = TRUE))
  knitr::kable(brfss_tb, digits = 2)

```


**5.For each year and state, compute the average proportion in each response category (taking the average across locations in a state). Make a five-panel plot that shows, for each response category separately, the distribution of these state-level averages over time.**

```{r problem 1.5}
aver_exce = 
  filter(brfss_df, response == "Excellent") %>%
  group_by(state,year) %>%
  summarize(mean_prop = mean(data_value)) %>%
  ggplot(aes(x = year, y = mean_prop, color = state)) +
  geom_line() +
  theme(legend.position = "none")

aver_very = 
  filter(brfss_df, response == "Very good") %>%
  group_by(state,year) %>%
  summarize(mean_prop = mean(data_value)) %>%
  ggplot(aes(x = year, y = mean_prop, color = state)) +
  geom_line() +
  theme(legend.position = "none")

aver_good = 
  filter(brfss_df, response == "Good") %>%
  group_by(state,year) %>%
  summarize(mean_prop = mean(data_value)) %>%
  ggplot(aes(x = year, y = mean_prop, color = state)) +
  geom_line() +
  theme(legend.position = "none")

aver_fair = 
  filter(brfss_df, response == "Fair") %>%
  group_by(state,year) %>%
  summarize(mean_prop = mean(data_value)) %>%
  ggplot(aes(x = year, y = mean_prop, color = state)) +
  geom_line() +
  theme(legend.position = "none")

aver_poor = 
  filter(brfss_df, response == "Poor") %>%
  group_by(state,year) %>%
  summarize(mean_prop = mean(data_value)) %>%
  ggplot(aes(x = year, y = mean_prop, color = state)) +
  geom_line() +
  theme(legend.position = "bottom")

aver_exce + aver_very + aver_good + aver_fair + aver_poor
  
```


## Problem 2 

**1. Import, explore and describe the dataset.**
```{r problem 2.2}
instacart_df = p8105.datasets::instacart 

